---
title: "Introduction to deduplication package"
author: "Bogdan Oancea"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette contains a short introduction to *deduplication* package. It describes its main purpose, presents some technical details of its implementation and provide examples on how to use this package. Some basic knowledge about *destim* package would be useful to understand how this package works. A detailed description of the methodological approach implemented by this package can be found in @WP5Deliverable1.3 and in @bmc_paper. To fully understand the theory behind this package it is recommended to read the above mentioned papers.


# Introduction
This section contains a brief explanation about the intended use of the package and provides a short introduction to the underlaying methodology.

## Device duplicity problem
The problem of device multiplicity comes from the fact that the statistical unit of analysis with mobile network data is the individual of a target population, not a mobile device. Since an individual can carry more than one device with him/her, this introduces the problem of multiple counting. For simplicity we make the simplifying assumption that each individual can carry at most $2$ devices. 

The main purpose of this package is to classify each device $d$ in a dataset as corresponding to an individual with only one device (1:1 correspondence between devices and individuals) or as corresponding to an individual with two devices (2:1 correspondence between devices and individuals). This classification will be probabilistic, thus assigning a probability $p_{d}$ of duplicity to each device $d$.

Two main methodological approaches are implemented by the *deduplication* package:

1. A Bayesian approach with network events data
2. An approach based on the distance between centers of location probabilities

## The Bayesian approach 
Denoting by $D_{d_i d_j}$ the event which has the meaning "devices $d_i$ and $d_j$ are carried by the same individual" and by$D^c_{d_i d_j}$ the event with the meaning "devices $d_i$ and $d_j$ are not carried by the same individual" the duplicity probability for device $d_i$ is the pair duplicity probability $p_{d_id_j} \equiv \mathbb{P}(D_{d_i d_j}| \mathbf{E}, \mathbf{I})$ corresponding to device $d_j$more similar to $d_i$ and we can write $p_{d_i}= \max_{d_j\neq d_i}\mathbb{P}\left(D_{d_id_j}|\mathbf{E}, \mathbf{I}\right)$. Obviously, $p_{d_i d_i} = 0$ for all $i$.

### The pair approach
We compute the pair-duplicity probabilities $p_{d_{i}d_{j}}$ for two devices $d_{i}$ and $d_{j}$ using the Bayes theorem: 

$$\mathbb{P}\left(D_{d_{i}d_{j}}^{c}|\mathbf{E}, \mathbf{I}\right)=\frac{\mathbb{P}\left(\mathbf{E}|D_{d_{i}d_{j}}^{c}\right)\mathbb{P}\left(D_{d_{i}d_{j}}^{c}| \mathbf{I}\right)}{\mathbb{P}\left(\mathbf{E}|D_{d_{i}d_{j}}\right)\mathbb{P}\left(D_{d_{i}d_{j}}| \mathbf{I}\right) + \mathbb{P}\left(\mathbf{E}|D_{d_{i}d_{j}}^{c}\mathbf{I}\right)\mathbb{P}\left(D_{d_{i}d_{j}}^{c}|\mathbf{I}\right)} =\frac{1}{1 + \frac{\mathbb{P}\left(D_{d_{i}d_{j}}|\mathbf{I}\right)}{\mathbb{P}\left(D_{d_{i}d_{j}}^{c}|\mathbf{I}\right)}\times \frac{\mathbb{P}\left(\mathbf{E}|D_{d_{i}d_{j}}, \mathbf{I}\right)}{\mathbb{P}\left(\mathbf{E}|D_{d_{i}d_{j}}^{c}, \mathbf{I}\right)} }$$

Here  $\mathbb{P}\left(D_{d_{i}d_{j}}|\mathbf{I}\right)$ and $\mathbb{P}\left(D_{d_{i}d_{j}}^{c}|\mathbf{I}\right)$ are the prior probabilities for the duplicity and non-duplicity events and

$\mathbb{P}\left(\mathbf{E}|D_{d_{i}d_{j}}, \mathbf{I}\right)$ and $\mathbb{P}\left(\mathbf{E}|D_{d_{i}d_{j}}^{c}, \mathbf{I}\right)$  stands for the likelihoods under each hypothesis $D_{d_{i}d_{j}}$ and $D_{d_{i}d_{j}}^{c}$, respectively.

After some mathematical manipulations we arrive at the following formula:

$$p_{d_{i}}=\max_{j\neq i}\left(\frac{1}{\left(1 + \alpha*\exp\left(\ell_{ij} - \ell_{i} -\ell_{j}\right)\right)}\right)$$, 

where $\alpha = \frac{P_2}{P_1}$, $P_1$ is the aprori probablity of duplicity for a device and $P_2=1-P_1$.

### The one-to-one approach
If we start from $p_{d_i} = \mathbb{P}\left( D_{d_i d_i} |  \mathbf{E}, \mathbf{I}  \right)$ and take into consideration  that the entire event set $\Omega_{d_i}$ for device $d_i$ can be decomposed as $\Omega_{d_i} = \cup_{d_j} D_{d_i d_j}$ we can apply again the Bayes theorem and obtain:

$$p_{d_i} = 1 - \frac{1}{1 + \alpha_{d_id_j} \times \sum_{j\neq i} \exp\ (\ell_i + \ell_j - \ell_{ij}) }$$

where $\ell_i$ and $\ell_j$ are loglikelihoods for single a device ($i$ and $j$ in this case), $\ell_{ij}$ is the loglikelihood for two the devices $d_i$ and $d_j$ and $\alpha = \frac{1-P_1}{P_1}$, $P_1$ is the apriori probability of 1:1 correspondence for all devices.

In both approaches, the likelihood for two devices $\ell_{ij}$ is computed using a similar HMM model as for one device with the difference in the emission model:  the emission probabilitiea are computed as as the product of the
original single-device emission probabilities for $d_i$ and $d_j$.

If we denote $\lambda^{(1)}_{d_i} = \frac{\mathbb{P}\left(D_{d_i d_i}|\mathbf{I}\right)}{1-\mathbb{P}\left(D_{d_i d_j}|\mathbf{I}\right)}$
the prior odds ratio which gives how much more probable is that an individual carries a priori only one device
$d_i$ than another device together with $d_i$, considering that apriori any other device $d_j$ can be the second device so that $\mathbb{P}\left( D_{d_i d_j}|\mathbf{I} \right)$ is constant for any other device $d_j$ and $\mathbb{P}\left( D_{d_id_i} | \mathbf{I} \right) + (N_D-1)* \mathbb{P}\left( D_{d_id_j} | \mathbf{I} \right) =1$, where $N_D$ is the total number of devices, we arrive at the following formula for the probability of duplicity:

$$p_{d_i}= 1 - \frac{1}{1+\frac{\exp\left(-\ell_{d_i}\right)}{\lambda_{d_i} \times \left(N_D-1\right)} \sum_{j \neq i} \exp\left(\ell_{d_i d_j}\right)}$$

Thus, depending on the available information, if there is information at the device level that can be used to evaluate $\lambda_{d_i}$, the latter formula can be used to compute the duplicity probability, otherwise, the former formula with a single value of $\alpha$ for all devices is used.

## The trajectory approach 

This approach also follows a Bayesian approach, but instead of using network event variables $\mathbf{E}$ we will use properties of the trajectories derived from the HMMs, i.e. the location probability distributions $\{\gamma_{d_{ti}}\}$ of all devices $d$.

Applying the Bayes theorem again we have:

$$\mathbb{P}\left(D_{d_{i}d_{j}}^{c}|\mathbf{X}, \mathbf{I}\right) = \frac{1}{1 + \frac{\mathbb{P}\left(D_{d_{i}d_{j}}|\mathbf{I}\right)}{\mathbb{P}\left(D_{d_{i}d_{j}}^{c}|\mathbf{I}\right)}\times    \frac{\mathbb{P}\left(\mathbf{X}|D_{d_{i}d_{j}}, \mathbf{I}\right)}{\mathbb{P}\left(\mathbf{X}|D_{d_{i}d_{j}}^{c}, \mathbf{I}\right)} }$$, 

where $\mathbf{X}$ is a variable related to the estimated trajectories in terms of posterior location probabilities.

To apply this idea we compute the probability distribution of the signed distance between the x- and y-axis position of each pair of devices (we denote these random variables by $\Delta_{x, d_i d_j t} = X_{d_i t} - X_{d_j t}$ and $\Delta_{y, d_i d_j t} = Y_{d_i t} - Y_{d_j t}$) and count how many times the mode of these distributions are less than a predefined quantity ($\gamma \times \max\left( r_{d_i t}, r_{d_j t} \right)$. If a device $d_{i}$ corresponds to an individual with two devices (2:1), there will be another device $d_{j}$ such that their distance will be significatively close to $0$ along their trajectories.

We define:

$$\hat{p}_{d_{i}d_{j}}^{\textrm{mode}}\equiv\mathbb{P}\left(\mathbf{X}|D_{d_{i}d_{j}}, \mathbf{I}\right) = \frac{\#\{t=1,\dots, T:|\delta_{xt}^{*}|\leq \xi\cdot\max\{rd_{d_{i}t}, rd_{d_{j}t}\}, |\delta_{yt}^{*}|\leq \xi\cdot\max\{rd_{d_{i}t}, rd_{d_{j}t}\}\}}{T}$$,

where $\delta_{xt}^{*}$ and $\delta_{yt}^{*}$ are the mode of the $\Delta_{x, d_i d_j t}$ and $\Delta_{y, d_i d_j t}$ variables. The duplicity probability will be given by:

$$p_{d_{i}} = \max_{j\neq i}\left(1 - \frac{1}{1 + \alpha\times \frac{\hat{p}_{d_{i}d_{j}}^{\textrm{mode}}}{1 - \hat{p}_{d_{i}d_{j}}^{\textrm{mode}}}}\right)$$

where $\alpha = \frac{P_1}{1-P_1}$, $P_1$ being the apriori probability of duplicity.

## Syntax and basic usage

This section explains briefly the main functions of the package and how to use each method of computing the duplicity probability implemented in the *deduplication* package.
In all the examples below we will use the data set included in this package. This data set was generated using the simulation software. The first step will be to set the path to the data:

```{r}
library(deduplication)
path_root <- 'extdata'
```

### The Bayesian approach with network events - pairs method
Next, we need a series of input parameters: 

```{r}
gridParams <-readGridParams(system.file(path_root, 'grid.csv', package = 'deduplication'))
```

*gridParams* is a list that contains the number of rows and columns of the grid and the tile dimensions along OX and OY axes.


Since we will use simulated data in our example, we need some parameters that were used to generate the data set: the probability of a person to have two mobile devices (needed for the apriori duplicity probability) and
the minimum value of the signal strength/quality to allow a connection between a mobile device and an antenna. These two values are read from the *simulation.xml* file:

```{r}
simParams <-readSimulationParams(system.file(path_root, 'simulation.xml', package = 'deduplication'))
```

We also need to read the network events:

```{r}
events <- readEvents(system.file(path_root, 'AntennaInfo_MNO_MNO1.csv', package = 'deduplication'))
```

From the *events* table we can easily build a list of devices present in the current data set and a table with the
antenna IDs where these devices are connected for every time instant:

```{r}
devices <- getDeviceIDs(events)
connections <- getConnections(events)
```

*devices* is a (sorted) list of the IDs of all devices detected by the network and *connections* is a matrix with
a each row corresponding to a device and the elements on the columns corresponding to the IDs of the antenna where 
the the device is connected at a time instant.

The next step will be the computation of the emission probabilities for the individual HMM models (for each device) and for the joint models (for pairs of devices). Since the emission probabilities are computed for each tile in the grid we need the number of rows and columns of the grid and also the file with the signal strength/quality. We also
need the minimum value of the signal strength/quality that allows a connection between a mobile device and an
antenna and we'll use it to set to 0 the signal strength/quality for each tile where the actual value is below
this threshold.

```{r}
emissionProbs <- getEmissionProbs(gridParams$nrow, gridParams$ncol, system.file(path_root, 'SignalMeasure_MNO1.csv', package = 'deduplication'), simParams$conn_threshold)

jointEmissionProbs <- getEmissionProbsJointModel(emissionProbs)
```

Using the emission probablities we can build the generic HMM model (for each individual device) and the joint HMM model (for pairs of devices)

```{r}
model <- getGenericModel(gridParams$nrow, gridParams$ncol, emissionProbs)
modelJ <- getJointModel(gridParams$nrow, gridParams$ncol, jointEmissionProbs)
```

#8. Fit models

ll <- fitModels(length(devices), model,connections)

#10. Read antenna cells and build a matrix of neighboring antennas

coverarea <- readCells(system.file(path_root, 'AntennaCells_MNO1.csv', package = 'deduplication'))

antennaNeigh <- antennaNeighbours(coverarea)

#11. Apriori probability of duplicity

P1 <- aprioriDuplicityProb(simParams$prob_sec_mobile_phone, length(devices))

#12. Build a matrix of pairs of devices to compute duplicity probability

pairs4dup<-computePairs(connections, length(devices), oneToOne = FALSE, P1=P1, limit = 0.05, antennaNeighbors = antennaNeigh)

#13. Compute duplicity probabilities using the "pairs" method (faster)

out1 <- computeDuplicityBayesian("pairs", devices, pairs4dup, modelJ, ll, P1)

#14. Apriori probability of 2-to-1

Pii <- aprioriOneDeviceProb(simParams$prob_sec_mobile_phone, length(devices))

#15. Build a matrix of pairs of devices to compute duplicity probability

pairs4dup<-computePairs(connections, length(devices), oneToOne = TRUE)

#16. Compute duplicity probabilities using "1to1" method

out2 <- computeDuplicityBayesian("1to1", devices, pairs4dup, modelJ, ll, P1 = NULL, Pii=Pii)

